---
title: "Dach1 scatter plot"
output:
  html_notebook:
    code_folding: hide
    toc: true
    toc_float: true
    collapsed: false
    fig_height: 15
    fig_width: 20
---

```{r}
markers_for_heatmap <- function(markers) {
    res <- NULL
    for (i in unique(markers[, 1])) {
        tmp <- markers[markers[, 1] == i, ]
        if (nrow(tmp) > 10) {
            res <- rbind(res, tmp[1:10, ])
        } else {
            res <- rbind(res, tmp)
        }
    }
    
    return(res)
}
organise_marker_genes <- function(object, k, p_val, auroc) {
    dat <- rowData(object)[, c(paste0("sc3_", k, "_markers_clusts"), paste0("sc3_", k, 
        "_markers_auroc"), paste0("sc3_", k, "_markers_padj"), "feature_symbol")]
    dat <- dat[dat[, paste0("sc3_", k, "_markers_padj")] < p_val & !is.na(dat[, paste0("sc3_", 
        k, "_markers_padj")]), ]
    dat <- dat[dat[, paste0("sc3_", k, "_markers_auroc")] > auroc, ]
    
    d <- NULL
    
    for (i in sort(unique(dat[, paste0("sc3_", k, "_markers_clusts")]))) {
        tmp <- dat[dat[, paste0("sc3_", k, "_markers_clusts")] == i, ]
        tmp <- tmp[order(tmp[, paste0("sc3_", k, "_markers_auroc")], decreasing = TRUE), ]
        d <- rbind(d, tmp)
    }
    
    return(d)
}

make_col_ann_for_heatmaps <- function(object, show_pdata) {
    if (any(!show_pdata %in% colnames(colData(object)))) {
        show_pdata_excl <- show_pdata[!show_pdata %in% colnames(colData(object))]
        show_pdata <- show_pdata[show_pdata %in% colnames(colData(object))]
        message(paste0("Provided columns '", paste(show_pdata_excl, collapse = "', '"), "' do not exist in the phenoData table!"))
        if (length(show_pdata) == 0) {
            return(NULL)
        }
    }
    ann <- NULL
    if (is.null(metadata(object)$sc3$svm_train_inds)) {
        ann <- colData(object)[, colnames(colData(object)) %in% show_pdata]
    } else {
        ann <- colData(object)[metadata(object)$sc3$svm_train_inds, colnames(colData(object)) %in% 
            show_pdata]
    }
    # remove columns with 1 value only
    if (length(show_pdata) > 1) {
        keep <- unlist(lapply(ann, function(x) {
            length(unique(x))
        })) > 1
        if (!all(keep)) {
            message(paste0("Columns '", paste(names(keep)[!keep], collapse = "', '"), "' were excluded from annotation since they contained only a single value."))
        }
        ann <- ann[, names(keep)[keep]]
        if (ncol(ann) == 0) {
            ann <- NULL
        } else {
            ann <- as.data.frame(lapply(ann, function(x) {
                if (nlevels(as.factor(x)) > 9) 
                  x else as.factor(x)
            }))
            # convert outlier scores back to numeric
            for (i in grep("_log2_outlier_score", colnames(ann))) {
                if (class(ann[, i]) == "factor") {
                  ann[, i] <- as.numeric(levels(ann[, i]))[ann[, i]]
                }
            }
        }
    } else {
        if (length(unique(ann)) > 1) {
            ann <- as.data.frame(ann)
            colnames(ann) <- show_pdata
            if (!grepl("_log2_outlier_score", show_pdata)) {
                ann <- as.data.frame(lapply(ann, function(x) {
                  if (nlevels(as.factor(x)) > 9) 
                    return(x) else return(as.factor(x))
                }))
            }
        } else {
            message(paste0("Column '", show_pdata, "' was excluded from annotation since they contained only a single value."))
            ann <- NULL
        }
    }
    return(ann)
}

sc3_plot_markers.new <- function(object, k, auroc, p.val, show_pdata) {
    if (is.null(metadata(object)$sc3$consensus)) {
        warning(paste0("Please run sc3_consensus() first!"))
        return(object)
    }
    hc <- metadata(object)$sc3$consensus[[as.character(k)]]$hc
    dataset <- get_processed_dataset(object)
    if (!is.null(metadata(object)$sc3$svm_train_inds)) {
        dataset <- dataset[, metadata(object)$sc3$svm_train_inds]
    }
    
    add_ann_col <- FALSE
    ann <- NULL
    if (!is.null(show_pdata)) {
        ann <- make_col_ann_for_heatmaps(object, show_pdata)
        if (!is.null(ann)) {
            add_ann_col <- TRUE
            # make same names for the annotation table
            rownames(ann) <- colnames(dataset)
        }
    }
    ann_colors = rep(list(c("grey","white", "firebrick2")),ncol(ann)-sum(sapply(ann, is.factor)))
    names(ann_colors) = colnames(ann)[!(sapply(ann, is.factor))]
    
    # get all marker genes
    markers <- organise_marker_genes(object, k, p.val, auroc)
    # get top 10 marker genes of each cluster
    markers <- markers_for_heatmap(markers)
    
    row.ann <- data.frame(Cluster = factor(markers[, 1], levels = unique(markers[, 1])))
    rownames(row.ann) <- markers$feature_symbol
    
    do.call(pheatmap::pheatmap, c(list(dataset[markers$feature_symbol, , drop = FALSE], show_colnames = FALSE, 
        cluster_rows = FALSE, cluster_cols = hc, cutree_cols = k, annotation_row = row.ann, annotation_colors = ann_colors, annotation_names_row = FALSE, cellheight = 10), list(annotation_col = ann)[add_ann_col]))
}
```


load the data
```{r}
source("sc_func.R")
suppressMessages(library(Rtsne))
suppressMessages(library(destiny))
suppressMessages(library(ggplot2))
suppressMessages(library(gplots))
suppressMessages(library(RColorBrewer))
suppressMessages(library(plotly))
suppressMessages(library(gridExtra))


hmcol<-colorRampPalette(c("blue", "white","red"))(n = 299)

load("sc_192.RData")
```
##heatmap
do PCA+diffusionmap and extract genes correlated with top PC, excepct PCs that related to cell cycles:

```{r}
pca_out = prcomp(t(log2(expr_data+1)))
dif <- DiffusionMap( pca_out$x[,c(2,3,5,7,8,9,10,11,12,13,14)],n_eigs = 10)
dif_GO_list = Diff_GO_enrichment(dif,expr_data,p.val=0.001)

diff_genes = c()
for(i in 1:10){diff_genes = c(diff_genes,rownames(dif_GO_list$DE_genes[[i]]))}
DE_expr = expr_data[rownames(expr_data) %in% diff_genes,]
DE_clean = DE_expr[apply(log2(DE_expr+1),1,mean)<6,]
DE_clean = DE_clean[rowMeans(DE_clean)>0.6,]
dim(DE_expr)


hr <- hclust(dist(cor(log2(DE_clean+1))))

tsne_ph = Rtsne(log10(ph_data[,!(colnames(ph_data) %in% c("CD11b"))]),pca = FALSE,perplexity = 10)
tsne_ph1 = Rtsne(log10(ph_data[,!(colnames(ph_data) %in% c("CD11b"))]),dim=1,pca = FALSE,perplexity = 10)

# define some clusters
mycl <- cutree(hr, k=10)
table(mycl)
#mycl[mycl>8] = 8 # exclude rare cluster
# get a color palette equal to the number of clusters
#clusterCols <- rainbow(length(unique(mycl)))
clusterCols = brewer.pal(10, "Set3")
# create vector of colors for side bar
myClusterSideBar <- clusterCols[mycl]
#ph_data$cluster = myClusterSideBar
DE_hm = heatmap.2(log2(DE_clean[,order(mycl)]+1),
          Colv=FALSE,
          trace="none",
          dendrogram = "none",
          col=hmcol,
          density.info = "none",
          labCol = "",
          labRow = "",
          ColSideColors = myClusterSideBar[order(mycl)],
          key = FALSE
          )

ma = log2(DE_clean[,order(mycl)]+1)
ma = ma[DE_hm$rowInd,]
plot_ly(
    x = colnames(ma), y = rownames(ma),
    z = ma, type = "heatmap",colorscale = hmcol
)
```


```{r}
database = "immgen"
immgen_samples <- read.delim(paste0("/Users/tian.l/Dropbox/research/Dach1_paper/ref_data/",database,"_samples.txt"), stringsAsFactors=FALSE)
immgen_probes <- read.delim(paste0("/Users/tian.l/Dropbox/research/Dach1_paper/ref_data/",database,"_probes.txt"), header=FALSE, stringsAsFactors=FALSE)
immgen_expression <- read.delim(paste0("/Users/tian.l/Dropbox/research/Dach1_paper/ref_data/",database,"_expression.txt"), stringsAsFactors=FALSE)

library('biomaRt')
mart <- useDataset("mmusculus_gene_ensembl", useMart("ensembl"))
G_list <- getBM(filters= "ensembl_gene_id", attributes= c("ensembl_gene_id", "external_gene_name", "description"),values=immgen_probes$V2 ,mart=mart)
immgen_probes = immgen_probes[immgen_probes$V2 %in% G_list$ensembl_gene_id,]
G_list = G_list[match(immgen_probes$V2, G_list$ensembl_gene_id),]
immgen_probes$external_gene_name = G_list$external_gene_name

immgen_expression = immgen_expression[immgen_expression$probeId %in% immgen_probes$V1,]
immgen_probes = immgen_probes[immgen_probes$V1 %in% immgen_expression$probeId,]
immgen_probes = immgen_probes[match(immgen_expression$probeId, immgen_probes$V1),]
immgen_expression$gene_id = immgen_probes$external_gene_name
immgen_expression = immgen_expression[!duplicated(immgen_expression$gene_id),]
g_i = immgen_expression$gene_id
immgen_expression = immgen_expression[,!(colnames(immgen_expression) %in% c("probeId","gene_id"))]
immgen_expression = as.matrix(immgen_expression)
rownames(immgen_expression) = g_i

table(immgen_samples$cell_lineage)
immgen_samples$sampleId = gsub("-",".",immgen_samples$sampleId)
#SC_list = immgen_samples[immgen_samples$cell_lineage == "Stem Cells", "sampleId"]
#SC_list = immgen_samples$sampleId
SC_list = c("SC_CDP_BM","SC_CMP_BM","SC_GMP_BM",
            "SC_LT34F_BM","SC_LTSL_BM","SC_MDP_BM",
            "SC_MEP_BM","SC_MPP34F_BM","SC_ST34F_BM",
            "SC_STSL_BM","proB_CLP_BM","preT_ETP_Th")
sub_immgen_expression = immgen_expression[rownames(immgen_expression) %in% rownames(DE_expr), colnames(immgen_expression) %in% SC_list]
sub_immgen_expression = sub_immgen_expression[!duplicated(rownames(sub_immgen_expression)),]
sub_DE_expr = DE_expr[rownames(DE_expr) %in% rownames(sub_immgen_expression),]
sub_DE_expr = sub_DE_expr[!duplicated(rownames(sub_DE_expr)),]
sub_DE_expr = sub_DE_expr[order(rownames(sub_DE_expr)),]
sub_immgen_expression = sub_immgen_expression[order(rownames(sub_immgen_expression)),]


p.val_mat = c()
for (i in 1:ncol(sub_immgen_expression))
{
  p.val_vec = apply(sub_DE_expr, 2, function(x){
  cor.test(x,sub_immgen_expression[,i],
         method = "spearman",
         alternative = "greater",exact=FALSE)$p.value})
  p.val_mat = rbind(p.val_mat,-log10(p.val_vec))
}
rownames(p.val_mat) = colnames(sub_immgen_expression)

hm_immgen = heatmap.2(p.val_mat[,order(mycl)],trace="none",
          dendrogram="none",
          Colv=FALSE,
          col=hmcol,
          scale="column",
          labRow = "",
          #ColSideColors = myClusterSideBar[order(mycl)],
          labCol = "",
          density.info="none",
          key=FALSE)

p_ma = p.val_mat[,order(mycl)]
p_ma = p_ma[hm_immgen$rowInd,]

m = list(
  l = 100,
  r = 40,
  b = 10,
  t = 10,
  pad = 0
) 

plot_ly(
    x = colnames(p_ma), y = rownames(p_ma),
    z = p_ma, type = "heatmap",colorscale = hmcol
)%>% 
layout(autosize = F, margin = m)
```

combine two heatmap:

```{r}
library(ComplexHeatmap)
ht1 = Heatmap(t(p.val_mat), name = "Immgen", show_row_names=FALSE, width = 1)
ht2 = Heatmap(t(log2(DE_clean+1)), name = "RNA-seq", show_row_names=FALSE, show_column_names=FALSE, width = 2)
ht3 = Heatmap(scale(log10(ph_data[,!(colnames(ph_data)=="CD11b")]+1)), name = "FACS",show_row_names=FALSE, width = 0.5)
ht_list = draw(ht3+ht2+ht1, main_heatmap = "RNA-seq", km = 10,show_row_dend=FALSE)

#clusterCols = brewer.pal(10, "Set3")
# create vector of colors for side bar
#myClusterSideBar <- clusterCols[mycl]

ht_list
```


##t-SNE

```{r}

tsne_diffmap_DE = Rtsne(cor(log2(DE_expr+1)))
tsne_diffmap_DE1 = Rtsne(cor(log2(DE_expr+1)),dim=1)
plot(tsne_diffmap_DE$Y[,1], tsne_diffmap_DE$Y[,2])

plot(tsne_diffmap_DE1$Y, tsne_ph1$Y)
```

##Dach1 correlation
```{r}
cor_shalin =cor(t(expr_data),method="spearman")

cor_Dach1 = cor_shalin["Dach1",]
cor_Dach1 = cor_Dach1[order(cor_Dach1)]
```


genes that positively correlated with Dach1:
```{r}
as.data.frame(head(cor_Dach1[order(cor_Dach1,decreasing=TRUE)],n=50))
```
some scatter plot:
```{r}
g_e = as.data.frame(t(log2(expr_data+1)))
p1=ggplot(data=g_e,aes(y=Dach1, x=Muc13))+geom_point()
p2=ggplot(data=g_e,aes(y=Dach1, x=Gata2))+geom_point()
p3=ggplot(data=g_e,aes(y=Dach1, x=Cdk17))+geom_point()
p4=ggplot(data=g_e,aes(y=Dach1, x=Sox6))+geom_point()

grid.arrange(p1, p2, p3, p4, ncol=2)
```

genes that negatively correlated with Dach1:


```{r}
as.data.frame(head(cor_Dach1[order(cor_Dach1)],n=50))
```

some scatter plot:

```{r}
p1=ggplot(data=g_e,aes(y=Dach1, x=Evl))+geom_point()
p2=ggplot(data=g_e,aes(y=Dach1, x=Il12a))+geom_point()
p3=ggplot(data=g_e,aes(y=Dach1, x=Dntt))+geom_point()
p4=ggplot(data=g_e,aes(y=Dach1, x=Flt3))+geom_point()

grid.arrange(p1, p2, p3, p4, ncol=2)
```
##Dach1 correlation in public data
data from Gottgens's Blood paper, we only select cells labeled as "HSC" and "LT-HSC".
```{r}
coordinates_gene_counts_flow_cytometry <- read.delim("/Users/tian.l/Dropbox/research/Dach1_paper/public_data/coordinates_gene_counts_flow_cytometry.txt", row.names=1, stringsAsFactors=FALSE)

coordinates_gene_counts_flow_cytometry = coordinates_gene_counts_flow_cytometry[complete.cases(coordinates_gene_counts_flow_cytometry),]
gene_expr = coordinates_gene_counts_flow_cytometry[,-(1:26)]
phen_data = coordinates_gene_counts_flow_cytometry[,6:15]

gene_expr_t = t(gene_expr)
G_cor = cor(t(gene_expr_t))

library('biomaRt')
mart <- useDataset("mmusculus_gene_ensembl", useMart("ensembl"))
G_list <- getBM(filters= "ensembl_gene_id", attributes= c("ensembl_gene_id", "external_gene_name", "description"),values=rownames(gene_expr_t) ,mart= mart)

gene_expr_t = gene_expr_t[rownames(gene_expr_t) %in% G_list$ensembl_gene_id,]
G_list = G_list[match(rownames(gene_expr_t), G_list$ensembl_gene_id),]
rownames(gene_expr_t) = G_list$external_gene_name

gene_expr_sub = gene_expr_t[,coordinates_gene_counts_flow_cytometry$group %in% c("HSC", "LT-HSC")]


corr_sub = cor(t(gene_expr_sub), method="spearman")

cor_Dach1_pub = corr_sub["Dach1",]
cor_Dach1_pub = cor_Dach1_pub[order(cor_Dach1_pub)]
```


genes that positively correlated with Dach1:
```{r}
as.data.frame(head(cor_Dach1_pub[order(cor_Dach1_pub,decreasing=TRUE)],n=50))
```

genes that negatively correlated with Dach1:


```{r}
as.data.frame(head(cor_Dach1_pub[order(cor_Dach1_pub)],n=50))
```
Genes that negatively correlated to Dach1 in both dataset (first 50):
```{r}
neg = toupper(names(head(cor_Dach1_pub[order(cor_Dach1_pub)],n=50)))
neg = neg[neg %in% toupper(names(head(cor_Dach1[order(cor_Dach1)],n=50)))]
neg
```

Genes that positively correlated to Dach1 in both dataset (first 50):
```{r}
pos = toupper(names(head(cor_Dach1_pub[order(cor_Dach1_pub,decreasing=TRUE)],n=50)))
pos = pos[pos %in% toupper(names(head(cor_Dach1[order(cor_Dach1,decreasing=TRUE)],n=50)))]
pos
```

##cell cycle

Cells are classified as being in G1 phase if the G1 score is above 0.5 and greater than the G2/M score; 
in G2/M phase if the G2/M score is above 0.5 and greater than the G1 score; 
and in S phase if neither score is above 0.5.

```{r}
library(scran)
mm.pairs <- readRDS(system.file("exdata", "mouse_cycle_markers.rds", package="scran")) 
library(org.Mm.eg.db) 
anno <- select(org.Mm.eg.db, keys=rownames(expr_data), keytype="SYMBOL", column="ENSEMBL") 
ensembl <- anno$ENSEMBL[match(rownames(expr_data), anno$SYMBOL)] 
assignments <- cyclone(log2(expr_data+1), mm.pairs, gene.names=ensembl) 
ggplot(data=NULL, aes(x=assignments$score$G1, y=assignments$score$G2M, col=log2(expr_data["Dach1",]+1)))+
  geom_point()+
  geom_vline(xintercept = .5)+
  geom_hline(yintercept = .5)


```


```{r}
cell_state = rep("S",ncol(expr_data))
cell_state[assignments$score$G1>0.5] = "G0/G1"
cell_state[assignments$score$G2M>0.5] = "G2/M"
dach1 = log2(expr_data["Dach1",]+1)
ggplot(data=NULL,aes(x=cell_state,y=dach1,fill=cell_state))+
  geom_violin(alpha=0.7) + geom_jitter(height = 0, width = 0.1)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank())
```

```{r}
public_assignments <- cyclone(t(gene_expr)[,coordinates_gene_counts_flow_cytometry$group %in% c("HSC", "LT-HSC")], mm.pairs, gene.names=rownames(t(gene_expr))) 
ggplot(data=NULL, aes(x=public_assignments$score$G1, y=public_assignments$score$G2M, col=gene_expr_t["Dach1",coordinates_gene_counts_flow_cytometry$group %in% c("HSC", "LT-HSC")]))+
  geom_point()+
  labs(x="G1",y="G2M",color="Dach1 expr")+
  geom_vline(xintercept = .5)+
  geom_hline(yintercept = .5)
```

```{r}
cell_state = rep("S",sum(coordinates_gene_counts_flow_cytometry$group %in% c("HSC", "LT-HSC")))
cell_state[public_assignments$score$G1>0.5] = "G0/G1"
cell_state[public_assignments$score$G2M>0.5] = "G2/M"
dach1 = gene_expr_t["Dach1",]
ggplot(data=NULL,aes(x=cell_state,y=dach1,fill=cell_state))+
  geom_violin(alpha=0.7) + geom_jitter(height = 0, width = 0.1)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank())
```



```{r}
library(SC3)
library(scater)
#sceset <- newSCESet(countData=DE_clean)
sceset = SingleCellExperiment(assays=list(counts=DE_expr[!duplicated(rownames(DE_expr)),],logcounts=log2(DE_expr[!duplicated(rownames(DE_expr)),]+1)))
rowData(sceset)$feature_symbol = rownames(sceset)
#sceset <- newSCESet(countData=DE_expr[!duplicated(rownames(DE_expr)),])
sceset = calculateQCMetrics(sceset)
sceset <- sc3(sceset, ks = 4:8, biology = TRUE,pct_dropout_max=95,rand_seed=19910603,kmeans_nstart=50)
sceset <- sc3_estimate_k(sceset)
```

```{r}

sceset$SC_LTSL_BM = p.val_mat["SC_LTSL_BM",]
sceset$SC_STSL_BM = p.val_mat["SC_STSL_BM",]
sceset$SC_MEP_BM = p.val_mat["SC_MEP_BM",]
sceset$SC_CMP_BM = p.val_mat["SC_CMP_BM",]
sceset$proB_CLP_BM = p.val_mat["proB_CLP_BM",]
sceset$SC_GMP_BM = p.val_mat["SC_GMP_BM",]

sceset$CD150 = log2(ph_data[, "CD150"]+1)
sceset$CD135 = log2(ph_data[, "CD135"]+1)
sceset$CD135[(sceset)$CD135<5] = 5
sceset$CD127 = log2(ph_data[, "CD127"]+1)
sceset$CD127[(sceset)$CD127<5] = 5
sceset$CD16.32 = log2(ph_data[, "CD16.32"]+1)
sceset$Sca1 = log2(ph_data[, "Sca.1"]+1)
sceset$cKit = log2(ph_data[, "cKit"]+1)
#sceset$G1 = assignments$scores$G1
#sceset$G2M = assignments$scores$G2M
```



```{r}
pdf("Dach1_clusters_IMMGEN.pdf",width = 8,height = 10)
sc3_plot_markers(sceset, k =  7,show_pdata = c(
        "SC_LTSL_BM", 
        "SC_STSL_BM",
        "SC_MEP_BM",
        "SC_CMP_BM",
        "proB_CLP_BM",
        "SC_GMP_BM",
        "sc3_7_clusters"
    ), p.val=0.01,auroc=0.75)
dev.off()
```


```{r}
pdf("Dach1_clusters_FACs.pdf",width = 8,height = 9)
sc3_plot_markers(sceset, k =  7,show_pdata = c(
        "CD150", 
        "CD135",
        "CD127",
        "CD16.32",
        "G1",
        "G2M"
    ), p.val=0.01,auroc=0.75)
dev.off()
```


```{r}
pdf("Dach1_clusters_all.pdf",width = 8,height = 16)
sc3_plot_markers.new(sceset, k =  5,show_pdata = c(
        "CD150", 
        "CD135",
        "CD127",
        "CD16.32",
        "SC_LTSL_BM", 
        "SC_STSL_BM",
        "SC_MEP_BM",
        "SC_CMP_BM",
        "proB_CLP_BM",
        "SC_GMP_BM"
    ), p.val=0.05,auroc=0.6)
dev.off()
```

```{r}
library(edgeR)


y = DGEList(counts=expr_data)
y = estimateCommonDisp(y)

design <- model.matrix(~log(expr_data["Dach1",]+1))
fit = glmFit(y,design=design)

lrt = glmLRT(fit,coef = 2)
t = topTags(lrt,n=Inf)

test_df = t@.Data[[1]]
test_df = test_df[!(rownames(test_df)=="Dach1"),]
test_df$FDR = -log10(test_df$FDR)

plot(test_df$logFC, test_df$FDR)

```

```{r}
test_df["Gata2",]
```


```{r}
library(scran)
sceall <- newSCESet(countData=expr_data[!duplicated(rownames(expr_data)),])
sceall = calculateQCMetrics(sceall)
var.fit <- trendVar(sceall, trend="loess", use.spikes=FALSE, span=0.2)
var.out <- decomposeVar(sceall, var.fit)
hvg.out <- var.out[order(var.out$bio, decreasing=TRUE)[1:2000],]
```


```{r}
scran_cor = correlatePairs(sceall, subset.row=rownames(hvg.out),BPPARAM=MulticoreParam(workers = 4))
save(scran_cor,file="/Users/tian.l/Dropbox/research/Dach1_paper/scran_cor_top2000.RData")
```

```{r}
scran_cor_dach1 = scran_cor[scran_cor$gene1 == "Dach1" | scran_cor$gene2 == "Dach1",]
the_cor_genes = rep(NA, nrow(scran_cor_dach1))
the_cor_genes[scran_cor_dach1$gene1 == "Dach1"] = scran_cor_dach1$gene2[scran_cor_dach1$gene1 == "Dach1"]
the_cor_genes[scran_cor_dach1$gene2 == "Dach1"] = scran_cor_dach1$gene1[scran_cor_dach1$gene2 == "Dach1"]

```


```{r}
library(readr)
renameME <- read_csv("~/Dropbox/research/Dach1_paper/renameME_LTMPP.csv")
renameME_DE = renameME[renameME$adj.P.Val<0.05,]
renameME_DE_CMP = renameME_DE[renameME_DE$FC>1,]
renameME_DE_CLP = renameME_DE[renameME_DE$FC<1,]

tmp = scran_cor_dach1$rho[scran_cor_dach1$FDR<0.05]

pdf("barcode_plot_Dach1_immgen_LT_MPP.pdf")
barcodeplot(scran_cor_dach1$rho, index=(the_cor_genes %in% renameME_DE_CMP$Gene_Symbol  ), 
            index2=(the_cor_genes %in%  renameME_DE_CLP$Gene_Symbol ), main="LTvsMPP34", 
            labels=c("negative correlation","positive correlation"),
            quantiles=c(max(tmp[tmp<0]),min(tmp[tmp>0])),
            xlab="speaman correlation to Dach1 expressions"
            )
dev.off()
```


```{r}
hm20 = colorRampPalette(c("blue", "white","red"))(n = 20)
side_bar = c("orchid1", "steelblue1")[as.factor(scran_cor_dach1$rho[scran_cor_dach1$FDR<0.05]>0)]
names(side_bar) = "rho to Dach1"
top_dach1_DE = the_cor_genes[scran_cor_dach1$FDR<0.05]
pdf("top_correlated_to_Dach1.pdf")
heatmap.2(cor(t(log2(expr_data[top_dach1_DE,]+1)),method = "spearman"),
          trace = "none",
          col=hmcol, 
          main="Correlation of genes correlated with Dach1",
          ColSideColors=side_bar,
          RowSideColors=side_bar,
          key.title="speaman rho")
legend("topright",      
    legend = c("positive correlated to Dach1", "negatively correlated to Dach1"),
    col = c("steelblue1","orchid1"),     
    lty= 1.5,   lwd = 2,           
    cex=.6)
dev.off()
```

```{r}
CODEX[CODEX$Factor %in% the_cor_genes,]
scran_cor_dach1[the_cor_genes %in% CODEX$Factor,]
```

```{r}
TF_genes = scPipe::get_genes_by_GO(returns="external_gene_name",
                            dataset="mmusculus_gene_ensembl",
                            go="GO:0003700")
```


```{r}
library(RColorBrewer)
darkcols <- brewer.pal(4, "Dark2")
sceset$cluster = sceset$sc3_5_clusters
pdf("FACS_overlay.pdf",width = 4,height = 3.5)
# cluster = as.factor(sceset$sc3_5_clusters == 1)
# ggplot(data=as.data.frame(colData(sceset)),aes(x=CD135,y=CD150,col=cluster))+
#   theme_bw()+
#   geom_point()+
#   scale_color_manual(values=c("grey80","red"))+
#   #labs(color="stem")+
#   ggtitle("stem")
#   
# cluster = as.factor(sceset$sc3_5_clusters %in% c(1,2))
# ggplot(data=as.data.frame(colData(sceset)),aes(x=CD135,y=CD150,col=cluster))+
#   theme_bw()+
#   geom_point()+
#   scale_color_manual(values=c("grey80","red"))+
#   #labs(color="stem+MEP")+
#   ggtitle("stem+MEP")
#   
# cluster = as.factor(sceset$sc3_5_clusters %in% c(1,2,3))
# ggplot(data=as.data.frame(colData(sceset)),aes(x=CD135,y=CD150,col=cluster))+
#   theme_bw()+
#   geom_point()+
#   scale_color_manual(values=c("grey80","red"))+
#   #labs(color="stem+MEP+myeloid")+
#   ggtitle("stem+MEP+myeloid")

ggplot(data=as.data.frame(colData(sceset))[sceset$cluster %in% c(4),],aes(x=CD150,y=CD135,col=cluster))+
  theme_minimal()+
  geom_point()+
  xlim(3.5, 10)+
  ylim(5, 13)+
  scale_color_manual(values=darkcols[c(4)])+
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("stem")

ggplot(data=as.data.frame(colData(sceset))[sceset$cluster %in% c(2,4),],aes(x=CD150,y=CD135,col=cluster))+
  theme_minimal()+
  geom_point()+
  xlim(3.5, 10)+
  ylim(5, 13)+
  scale_color_manual(values=darkcols[c(2,4)])+
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("stem+MEP")


ggplot(data=as.data.frame(colData(sceset))[sceset$cluster %in% c(2,3,4),],aes(x=CD150,y=CD135,col=cluster))+
  theme_minimal()+
  geom_point()+
  xlim(3.5, 10)+
  ylim(5, 13)+
  scale_color_manual(values=darkcols[c(2,3,4)])+
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("stem+MEP+myeloid")

ggplot(data=as.data.frame(colData(sceset)[sceset$cluster %in% c(1,2,3,4),]),aes(x=CD150,y=CD135,col=cluster))+
  theme_minimal()+
  geom_point()+
  xlim(3.5, 10)+
  ylim(5, 13)+
  scale_color_manual(values=darkcols)+
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("stem+MEP+myeloid+lymphoid")
dev.off()
```


```{r}
pdf("FACS_plot_per_clusters.pdf")
cluster = as.factor(pData(sceset)$sc3_7_clusters == 1)
ggplot(data=pData(sceset),aes(x=CD135,y=CD150,col=cluster))+geom_point()+ggtitle("cluster 1")
ggplot(data=pData(sceset),aes(x=Sca1,y=CD127,col=cluster))+geom_point()+ggtitle("cluster 1")
ggplot(data=pData(sceset),aes(x=Sca1,y=CD16.32,col=cluster))+geom_point()+ggtitle("cluster 1")

cluster = as.factor(pData(sceset)$sc3_7_clusters == 2)
ggplot(data=pData(sceset),aes(x=CD135,y=CD150,col=cluster))+geom_point()+ggtitle("cluster 2")
ggplot(data=pData(sceset),aes(x=Sca1,y=CD127,col=cluster))+geom_point()+ggtitle("cluster 2")
ggplot(data=pData(sceset),aes(x=Sca1,y=CD16.32,col=cluster))+geom_point()+ggtitle("cluster 2")

cluster = as.factor(pData(sceset)$sc3_7_clusters == 3)
ggplot(data=pData(sceset),aes(x=CD135,y=CD150,col=cluster))+geom_point()+ggtitle("cluster 3")
ggplot(data=pData(sceset),aes(x=Sca1,y=CD127,col=cluster))+geom_point()+ggtitle("cluster 3")
ggplot(data=pData(sceset),aes(x=Sca1,y=CD16.32,col=cluster))+geom_point()+ggtitle("cluster 3")

cluster = as.factor(pData(sceset)$sc3_7_clusters == 4)
ggplot(data=pData(sceset),aes(x=CD135,y=CD150,col=cluster))+geom_point()+ggtitle("cluster 4")
ggplot(data=pData(sceset),aes(x=Sca1,y=CD127,col=cluster))+geom_point()+ggtitle("cluster 4")
ggplot(data=pData(sceset),aes(x=Sca1,y=CD16.32,col=cluster))+geom_point()+ggtitle("cluster 4")

cluster = as.factor(pData(sceset)$sc3_7_clusters == 5)
ggplot(data=pData(sceset),aes(x=CD135,y=CD150,col=cluster))+geom_point()+ggtitle("cluster 5")
ggplot(data=pData(sceset),aes(x=Sca1,y=CD127,col=cluster))+geom_point()+ggtitle("cluster 5")
ggplot(data=pData(sceset),aes(x=Sca1,y=CD16.32,col=cluster))+geom_point()+ggtitle("cluster 5")

cluster = as.factor(pData(sceset)$sc3_7_clusters == 6)
ggplot(data=pData(sceset),aes(x=CD135,y=CD150,col=cluster))+geom_point()+ggtitle("cluster 6")
ggplot(data=pData(sceset),aes(x=Sca1,y=CD127,col=cluster))+geom_point()+ggtitle("cluster 6")
ggplot(data=pData(sceset),aes(x=Sca1,y=CD16.32,col=cluster))+geom_point()+ggtitle("cluster 6")

cluster = as.factor(pData(sceset)$sc3_7_clusters == 7)
ggplot(data=pData(sceset),aes(x=CD135,y=CD150,col=cluster))+geom_point()+ggtitle("cluster 7")
ggplot(data=pData(sceset),aes(x=Sca1,y=CD127,col=cluster))+geom_point()+ggtitle("cluster 7")
ggplot(data=pData(sceset),aes(x=Sca1,y=CD16.32,col=cluster))+geom_point()+ggtitle("cluster 7")
dev.off()
```


```{r}

y = DGEList(counts=expr_data)
y = estimateCommonDisp(y)

design <- model.matrix(~cell_state)
fit = glmFit(y,design=design)

lrt = glmLRT(fit,coef = 2)
cc_genes = rownames(topTags(lrt,n=Inf,p.value=0.05))
lrt = glmLRT(fit,coef = 3)
cc_genes = c(cc_genes, rownames(topTags(lrt,n=Inf,p.value=0.05)))
```


```{r}
sce_hi = sceall[rownames(hvg.out),]
#sce_hi = sce_hi[!(rownames(sce_hi) %in% cc_genes),]
```

```{r}
sce_hi <- sc3(sce_hi, ks = 4:10, biology = TRUE)
sce_hi <- sc3_estimate_k(sce_hi)
pData(sce_hi)$G1 = assignments$scores$G1
pData(sce_hi)$G2M = assignments$scores$G2M
```

```{r}
sc3_plot_markers(sce_hi, k =  7,show_pdata = c(
        "G1",
        "G2M"
    ), p.val=0.01,auroc=0.75)
```

## revisit t-SNE plots

```{r}
set.seed(42)
sc3_cor = sceset@sc3$consensus$`8`$consensus
sc3_cor = sc3_cor+matrix(0.00001*rnorm(nrow(sc3_cor)*ncol(sc3_cor)),nrow=nrow(sc3_cor),ncol=ncol(sc3_cor))
tsne_sc3 = Rtsne(1-sc3_cor,perplexity = 20,pca=FALSE,is_distance=TRUE)
plot(tsne_sc3$Y, col=as.factor(pData(sceset)$sc3_7_clusters))
```

```{r}
tsne_sc31 = Rtsne(sc3_cor,perplexity = 20,pca=FALSE)
plot(tsne_sc31$Y, col=as.factor(pData(sceset)$sc3_7_clusters))
```


```{r}
rtsne_ph = Rtsne(pData(sceset)[,c("CD16.32","CD135","CD127","CD150","Sca1","cKit")],perplexity = 20,pca=FALSE,dim=1)

tsne_sc3dim1 = Rtsne(1-sc3_cor,perplexity = 20,pca=FALSE,is_distance=TRUE,dims = 1)

ggplotly(ggplot(data=NULL,aes(x=tsne_sc3dim1$Y[,1],y=rtsne_ph$Y[,1],col=as.factor(pData(sceset)$sc3_8_clusters)))+geom_point())
```

```{r}
ggplotly(ggplot(data=NULL,aes(x=tsne_sc3$Y[,1],y=tsne_sc3$Y[,2],col=rtsne_ph$Y[,1]))+geom_point())
```

```{r}
ggplotly(ggplot(data=NULL,aes(x=tsne_sc3$Y[,1],y=tsne_sc3$Y[,2],col=exprs(sceset)["Dach1",]))+geom_point())
```


```{r}
ggplotly(ggplot(data=NULL,aes(x=tsne_sc3$Y[,1],y=tsne_sc3$Y[,2],col=as.factor(pData(sceset)$sc3_8_clusters)))+geom_point())
```


```{r}
#plot_ly(x=tsne_sc3$Y[,1],y=tsne_sc3$Y[,2],z=rtsne_ph$Y[,1],color=as.factor(pData(sceset)$sc3_8_clusters))
```



```{r}
plot_ly(x=tsne_sc3$Y[,1],y=rtsne_ph$Y[,1],color=as.factor(pData(sceset)$sc3_8_clusters))
```

```{r}
plot_ly(x=tsne_sc3$Y[,2],y=rtsne_ph$Y[,1],color=as.factor(pData(sceset)$sc3_8_clusters))

```

